---
title: "FANTOM OneGenE"
output: html_notebook
---
Descriptive statistics for the FANTOM OneGenE output data

Loading the matrix and printing stats
```{r message=FALSE}
library(data.table)
matrix = fread(
  file = "/home/silvio/onegene/FANTOM_unified.csv",
  sep = ';',
  header = T,
  verbose = F
)
min(abs(matrix$PearsonCorrelation))
summary(matrix$PearsonCorrelation)
summary((matrix$`relfreq-|pearson|`))
```
Plotting the relative frequencies
```{r}
summary(matrix$RelativeFrequency)

library(ggplot2)
ggplot(matrix, aes(x = RelativeFrequency)) + #Vedere se va il riferimento alla colonna
  scale_x_continuous(breaks = seq(0, 1, 0.05)) +
  geom_histogram(bins = 50) +
  theme_minimal() +
  labs(title = "Relative frequencies distribution",
       x = "Relative frequency",
       y = "Count")
gc(verbose = F)
```

Correlation distribution
```{r}
library(ggplot2)
plot = ggplot() +
  theme_minimal() +
  geom_density(aes(matrix$PearsonCorrelation)) +
  scale_x_continuous(breaks = seq(-0.75, 1, 0.1)) +
  geom_vline(aes(xintercept = mean(matrix$PearsonCorrelation)),
             color = "#0f4d92") +
  labs(
    x = "Pearson correlation coefficient",
    y = "Density",
    title = "Transcripts with relative frequency > 0",
    caption = "Blue line = mean"
  )
plot
```
Scatterplot of relative freq and correlation
```{r}
cor(matrix$PearsonCorrelation, matrix$RelativeFrequency)
#Useless clutter
library(scattermore)
scattermoreplot(
  matrix$RelativeFrequency,
  matrix$PearsonCorrelation
)
```
Scatterplot of relative freq and absolute value of correlation
```{r}
#Does not work
cor(abs(matrix$PearsonCorrelation), matrix$RelativeFrequency)

ggplot(matrix, aes(x=PearsonCorrelation, y=RelativeFrequency) ) +
  geom_density_2d()

# ggplot(matrix, aes(x=RelativeFrequency, y=abs(PearsonCorrelation))) +
#   stat_density_2d(aes(fill = ..level..), geom = "polygon", colour="white")
```
Number of expansions with corr>0.046
```{r}
nrow(matrix[abs(matrix$PearsonCorrelation) > 0.046,])
```
Loading pearson triangular matrix and counting corr>0.046
```{r}
#Exec on server because the matrix is 80 GB
library(data.table)
matrix = fread(file = "/storage/shared/fantom/FANTOM_PearsonMatrix_triangular.csv", sep = ',', header = T, verbose = F)
matrix = abs(matrix)
sum(matrix > 0.046, na.rm = T)
```

Violin + box plots
```{r}
library(ggplot2)
library(tidyr)

colnames(matrix)[3] = "Relative frequency"
colnames(matrix)[4] = "Pearson correlation"
colnames(matrix)[5] = "Rel freq - |corr|"
prova = matrix %>% pivot_longer(3:5)
prova$name = factor(prova$name,
                    levels = c("Relative frequency", "Pearson correlation", "Rel freq - |corr|"))

ggplot(prova, aes(name, value, color = name)) +
  scale_y_continuous(breaks = seq(-1, 1, 0.25)) +
  geom_violin() +
  geom_boxplot(width = 0.1) +
  theme_minimal() +
  scale_color_manual(values=c("#0f4d92", "#ff832f", "#a8bfd7"))+
  xlab("")+
  guides(color=F, fill=F)
```



